Symmetric Peer-to-Peer Applications

### 2024-10-17 rita-24/response.md

# Response Letter

We would like to thank the reviewers for the thoughtful comments, questions,
and criticism.

We accommodate all clarifications and answers in a final version of the paper.
To help the reviewers find the modifications of interest, we provide a link
with a *diff* between the versions, as follows:

https://

Based on the overall comments, our focus on the final version was to make the
limitations of our work more explicit, and to improve the section on related
work.

We also address each individual comment from all reviewers as follows...

- usar argumento de que centralizados precisam de bcast
- nosso caso é p2p, nem existe possibilidade de bcast pois nos nao tem
  conhecimento da rede
- possibilidade vem da existencia da time machine, que nao existe nos centralizados

In particular, the dependency on a central server

- target small P2P networks
    - examples / topology extrapolate the target

- pq exec local?
we use an absolute simulation clock to track

The topology has an average of 5 hops between peers
    - ja é bastante radical

We use NetEm [9] to simulate the network latency with
a normal distribution.

The experiments take around 40 hours to complete.

40s online and 20s offline in the
average, including peers 9-11 in the middle of the network.

## Reviewer A

Regarding the experimental evaluation:

> <...>
> However, the experimental evaluation could be more extensive and rigorous.
> <...>
> The proposed approach is technically sound and the experimental results are
> promising.
> <...>
> b) The experimental evaluation could be more extensive, including more
> diverse scenarios and a comparison with existing approaches
> <...>
> The evaluation uses a fixed network topology. How would the middleware's
> performance be affected in scenarios with highly dynamic topologies, where
> peers frequently join, leave, or experience intermittent connectivity?
> Have you considered any mechanisms to adapt to such changes?
> <...>
> The evaluation focuses on simulations with up to 21 peers. How do you
> envision the middleware scaling to much larger networks, and what challenges
> might arise in real-world deployments with potentially hundreds or thousands
> of peers?

Regarding the section on related work:

> a) The related work section could be expanded to include more recent work on
> peer-to-peer systems and distributed consensus algorithms;

Regarding the limiations:

> <...>
> c) The paper could discuss the limitations of the proposed approach and
> potential future work.
> <...>
> The paper assumes non-malicious peers. What security measures or
> modifications would be necessary to protect against malicious peers
> attempting to disrupt the synchronization or inject false events?

Regarding the middleware & protocol:

> While the API description is generally clear, some parts could be further
> clarified, particularly regarding the handling of dynamic memory allocation
> and the interaction between the application and the middleware during time
> travel.
> <...>
> Suggestion: if possible please consider providing a more formal description
> of the time synchronization protocol, including any guarantees or assumptions
> about clock drift and network delays.

## Reviewer J

Regarding the section on related work (comments J.1-J.6):

### Comment J.1

> i) This can do with a subtitles to make it easier to follow the discussion.
> There appears to be two major parts in this section: (1) the discussion on
> the three selected works (Croquet, GALS, CRDTs) in comparison to this work
> and this Work and (2) related works on time machines. It would be beneficial
> to clearly delineate these two subsection, preferably with subtitles.

As suggested, we split Section 2 in two:

- 2.1. Symmetric Distributed Applications
- 2.2. Software Time Machines

### Comment J.2

> ii) The introduction on the related works makes it appear like there are
> ideally 3 subsections: (a) how the network is organized, (b) how global time
> is determined, and (c) how events are propagated and applied.
> However, this three aspects are actually related to the discussion on the
> selected works compared to the proposed solution.

To make clear that we have only two subsections, we now briefly introduce 2.1
and 2.2 with paragraph at the beginning of 2 (before 2.1 and 2.2):

### Comment J.3

> This section first revisits existing solutions for symmetric distributed
> applications, namely Croquet [2], GALS [1], and CRDTs [3]. Then, we discuss
> software time machines that are built for either local or networked
> applications.

We now link the three aspects (a,b,c) with the selected works at the beginning
of 2.1:

> Even though existing solutions aim to provide symmetric behavior for
> distributed applications, they diverge in three key aspects as follows:
> a. How the network is organized.
> b. How global time is determined.
> c. How events are propagated and applied.
> Figure 1 compares three selected works regarding these aspects.
> <...>

### Comment J.4

> --> There is a need to reorganize the related works section, with a proper
> introduction showing the two core aspects of the discussion, (1) and (2) as
> indicated in i), forming the introduction to the section.

Done.

### Comment J.5

> iii) The discussion does not have a very smooth transition. The punchline of
> the related works section should ideally be the proposal of this work, which
> is brought out more clearly when discussed at the end of the related works to
> show how all the gaps identified will be solved.

To guide our discussion, we explain how the text is organized at the beginning
of the section:

> <...>
> Figure 1 compares three selected works regarding these aspects. We now
> discuss these works with respect to the three aspects, while comparing with
> our approach. At the end of this section, we also provide initial details on
> how our middleware works.

### Comment J.6

> iv) The discussion on related work on time machines is rather abrupt.
> It does not have a proper introduction to the problem of time machines in
> relation to this work (or in general for that matter) but directly proceeds
> into an analysis of Fusion and Braid.
> --> What is aimed by this section? Answering this can help find a way of
> introducing the time machines aspect well to the reader.

It was indeed abrupt and out of context.
We now start discussing (a) what a software time machine is, (b) what are the
difficulties to xxx, and (c) what are the common implementation techniques.
Then, we

Regarding peer-to-peer networks, applications, and protocols (comments J.1-J.6):

### Comment J.7

> <...>
> However, in a discussion on p2p application, or even systems, it is necessary
> to highlight the underlying infrastructure to dispel any doubts.
> In this regard, the reviewer would have expected to see the following
> addressed:

> i) As a proof-of-concept, it will be important to have p2p
> model/framework to give a clear picture to an expert reader and explain to a
> lay reader the place of this middleware in the overall p2p application
> model/framework. Has this been considered?

> ii) This proposal uses broadcasting to synchronize peers. What is the
> implication of this assuming a large p2p network? Has this been considered?

> iii) How are the events, as well as the memory snapshots, stored for
> persistence in the p2p network, that is, is there a storage mechanisms used
> so that an offline who joins immediately picks up from where it had dropped off?

> iv) The simulations were done on the same machine. This means that there is
> no actual considerations due to the effect of network challenges. Is this the
> case?

> v) Although this was considering the middleware only, it would be important
> to talk about the impact in terms of network efficiency if any due to the
> middleware.

Regarding grammar issues:

### Comment J.X

> Figure 2 - GALS --> check the spelling for client

OK.

> <...>
> This should be ...among all peers to be the “correct real time”,...

OK.

### 2024-12-06 rita-24/rita.tex
- versao final

\documentclass[sn-mathphys,iicol]{sn-jnl}

\jyear{2024}

\raggedbottom

\usepackage{verbatim}
\usepackage{url}
\usepackage{graphicx}

\usepackage{xspace}
\newcommand{\FC}       {Freechains\xspace}
\newcommand{\reps}     {\emph{reps}\xspace}
\newcommand{\onerep}   {\emph{1~rep}\xspace}
\newcommand{\nreps}[1] {\emph{#1~reps\xspace}}
\newcommand{\code}[1]  {\texttt{\footnotesize{#1}}}
\newcommand{\Xon} {$1{\rightarrow}N$\xspace}
\newcommand{\Xno} {$1{\leftarrow}N$\xspace}
\newcommand{\Xnn} {$N{\leftrightarrow}N$\xspace}
\newcommand{\Xoo} {$1{\leftrightarrow}1$\xspace}
\newcommand{\Xo}  {$1{\hookleftarrow}$\xspace}
\newcommand{\lin}[1]{(\emph{ln. #1}\xspace)}
\newcommand{\linx}[1]{\emph{ln. #1}\xspace}

\renewcommand{\theenumi}{\alph{enumi}}

\hyphenation{off-line}

@@

\title[Article Title]{
    Symmetric Peer-to-Peer Applications
}

\author{\fnm{Francisco} \sur{Sant'Anna} \email{francisco@ime.uerj.br}}
\affil{\orgdiv{Department of Computer Science}, \orgname{Rio de Janeiro State University}, \orgaddress{\country{Brazil}}}

\abstract{
In real-time networked applications, such as shared documents, users can
interact remotely and yet share the exact same experience as if they were in a
single machine.
%
In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications} in which decentralized instances can broadcast asynchronous
events and yet conform to identical behavior.
%
Peers are allowed to join and leave the network at any time.
Application developers must adhere to a restricted API, which is deterministic,
stateless, and only supports pre-allocated memory.
%
The middleware is responsible for synchronizing the events in a global shared
timeline across the network.
Also, based on memory snapshots and deterministic simulation, a ``time machine''
can rollback conflicting peers to resynchronize their state.
%
We show that the middleware can handle applications with over 20 peers in a
heterogeneous topology under high churn.
For instance, in a simulation scenario of 25 events per minute with a network
latency of 50ms, the peers need to roll back and resynchronize only 3\% of
the time.
}

\keywords{
collaboration, determinism, peer-to-peer, time machine
}

\maketitle

\section{Introduction}
\label{sec.introduction}

@@

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Peers in the application form a dynamic network graph and communicate only
with direct neighbours, as in typical unstructured peer-to-peer networks.
Events are flooded in the graph and are triggered locally with a small delay
to compensate the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.
Our main contribution is the design of a middleware to ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

@@

In Section~\ref{sec.related}, we revisit existing solutions for symmetric
distributed applications.
In Section~\ref{sec.tml}, we detail the architecture of our middleware, its
programming API, and how it orchestrates peers.
%In Section~\ref{sec.app}, ...
In Section~\ref{sec.eval}, we evaluate our design under a number of scenarios.
In Section~\ref{sec.conclusion}, we conclude this work.

@@

@@

In this section, we revisit existing solutions for symmetric distributed
applications.
We focus on
    (a) how the network is organized,
    (b) how global time is determined, and
    (c) how events are propagated and applied.
Figure~\ref{fig.table} compares three selected works regarding these aspects.

Croquet%
\footnote{Croquet.io: https://croquet.io/}~\cite{croquet} guarantees
bit-identical real-time behavior for users in collaborative distributed
environments.
%
As detailed in Figure~\ref{fig.algos}, a centralized \emph{reflector} issues
periodic ticks, such that all nodes in the network advance together according
to a synchronized global shared clock.
If a user generates an event, it is sent back to the reflector, which
broadcasts the event in the next tick, which all nodes apply in sync.
%
Croquet takes periodic snapshots of the whole application state in order to
support late joins to a running session.
The new node just needs to restore the latest snapshot and simulate the
remaining events to reach the current running state.
%
As indicated in Figure~\ref{fig.table}, Croquet relies on a centralized
network (a), in which nodes advance in sync (b), and in which event outcomes
depend on the central server to be online (c).

GALS~\cite{gals} shares the same goals with Croquet, but with some tradeoffs,
mostly favoring clients with slow connections.
As detailed in Figure~\ref{fig.algos}, instead of advancing ticks in sync with
the server, clients have their own independent local clocks.
Event generation requires two round trips (\emph{when→ask→ack→tick}) and is
delayed dynamically according to the slowest client.
%
On the one hand, ticks do not generate any traffic and clients have smooth
frame transitions, even those with poor connections.
On the other hand, events take longer to apply and clients may experience
occasional freezes.
The syncing protocol also requires extra bookkeeping to deal with clock drifts
and client disconnections~\cite{gals}.
%
As indicated in Figure~\ref{fig.table}, GALS also relies on a centralized
network (a), but in which nodes advance time independently (b), and in which
event outcomes still depend on the central server to be online (c).

@@

In this work, as indicated in Figure~\ref{fig.table}, our goal is to support
peer-to-peer real-time symmetric execution (a), while still tolerating
out-of-order (b) and offline (c) event generation.
%
As detailed in Figure~\ref{fig.algos}, peers have independent timelines and
can communicate events directly to each other.
Events are delayed to be able to reach other peers in time.
In the case a deadline is missed, the peer rolls back in time and then fast
forwards execution until it catches real-time behavior again.
%
Like Croquet and GALS, peers compute events in sequence using deterministic and
stateless calls only.

We now discuss related work on time machines.
%
Regarding decentralized machines, Fusion%
\footnote{Fusion: \url{https://doc.photonengine.com/en-us/fusion}}
is a state synchronization networking library for the Unity game engine.
It provides networked tick-based simulation in two modes: client-server
(\emph{hosted mode}) and peer-to-peer (\emph{shared mode}).
The hosted mode is based on continuous memory snapshots, which allows for
full rollbacks when clients diverge.
However, the shared mode is less powerful and only supports eventual
consistency among clients.
We presume that continuous memory snapshots would be too costly without a
central server.
%
Regarding single-user local machines, the video game Braid%
\footnote{Braid: \url{http://braid-game.com/}}
is designed on the assumption that players have unrestricted rollbacks as part
of the game mechanics.
Like Fusion, the implementation is based on continuous memory snapshots,
instead of deterministic simulation as we propose in this work.
%
Schuster~\cite{tml.js} proposes a simple API for time travelling in
event-driven JavaScript:
    \code{init()} to return an initial state;
    \code{handle(evt,mem)} to process events and modify state; and
    \code{render(mem)} to output the current state.
As we describe next, we use a similar approach by separating state modification
from rendering~\cite{tml.alive}.

@@

We guide our discussion through the didactic example of Figure~\ref{fig.move},
in which multiple users use the arrow keys to control the direction of a
moving rectangle.
The six instances are connected in a peer-to-peer network, such that most
peers can only communicate indirectly to each other.
%
The application is symmetric in the sense that if any user presses a key, a
corresponding event is broadcast and applied in all peers simultaneously, as
if they were a single mirrored machine.
At any moment, if a user pauses the application, all peers are guaranteed to
pause exactly at the same frame with the rectangle being ``bit-identically'' at
the same position.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{middleware}
  \caption{
    Applications generate asynchronous events and communicate through an API
    with the middleware.
    The middleware orchestrates the peers, synchronizing their ticks locally
    and broadcasting events with a $\Delta$ delay.
    \label{fig.middleware}
  }
\end{figure}

In Figure~\ref{fig.middleware}, a peer is depicted as an application that uses
an API to communicate with the middleware.
The application itself is not aware of the network and has no direct access to
other peers, which communicate transparently through the middleware.
The application generates asynchronous events that go through the middleware,
which synchronizes them with a timestamp with a $\Delta$ delay scheduled to the
future, such that all peers can satisfy.
The middleware controls the execution of the application by issuing ticks and
synchronized events.

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
As described in the Introduction, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate $\Delta$ so that they can reach other peers in time to be applied in
sync.
%We first assume a predetermined delay for the sake of simplicity.
Nevertheless, a peer ahead of time may receive an event that should have been
applied in the past, even considering the delay.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
As we detail further, rolling back a peer requires to simulate the whole
application execution from the beginning.
As an optimization, the middleware takes periodic snapshots locally to avoid
full simulation.
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because applications are launched locally at different times, and
because internal clocks may diverge over time (e.g., from rollbacks or timer
polling inaccuracies).
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
\item \textbf{Peer Churn:}
Considering that we target peer-to-peer networks, it is important that the
middleware can handle arrival and departure of peers seamlessly.
Note that high churn may lead to intermittent network partitions.
Given our unstructured approach, nothing needs to be done when a peer leaves
the network.
However, when a peer joins the network, the middleware needs to ensure that it
receives from other peer all events that ever happened, in order.
This is trivial since all peers already need to hold the full event history (as
described in item (a)), including their absolute timestamps in the shared
timeline.
\end{enumerate}

@@

When an application generates events, they need to be broadcast to the network
so that all peers behave the same.
Note that application events may (or may not) differ from low-level local
system events.
%which may or may not be mapped 1-to-1 in the callback \code{cb\_evt}.
%
The middleware predefines two events for all applications:
    \code{P2P\_EVT\_START} represents the beginning of the simulation, and
    \code{P2P\_EVT\_TICK} is generated every frame.
The other application-specific events must be declared by the programmer in an
enumeration.
In our example, we define a new event \code{P2P\_EVT\_KEY} to represent key
presses:

@@

The callback \code{cb\_evt} is called by the middleware in real time, whenever
a local SDL event occurs.
This callback has two goals:
    (a) map and broadcast application events, and
    (b) provide instant feedback to the user.
%As mentioned previously,
Not all local low-level events need to broadcast application events in the
network.
The callback is allowed to perform side effects, such as modifying the network
topology, or exhibiting instant feedback on the screen.
In our example, we only generate application events for the 5 key events of
interest, but we also exhibit the pressed key in real time as a visual
feedback:

@@

\subsubsection{Event Broadcasting}
\label{sec.tml.middleware.events}

>>>>

\documentclass[fleqn,10pt]{SelfArx}
\usepackage[english]{babel}
\usepackage{lipsum}
\usepackage{url}
\usepackage{balance}

\newcommand{\code}[1]  {\texttt{\footnotesize{#1}}}
\newcommand{\lin}[1]{(\emph{ln. #1}\xspace)}
\newcommand{\linx}[1]{\emph{ln. #1}\xspace}

\usepackage{xspace}
\usepackage{hyperref}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{textcomp}
\usepackage{gensymb} % degree. ie 360º
\usepackage{amsmath}
\usepackage{comment}
\usepackage{multirow}
\usepackage[framemethod=tikz]{mdframed}
\usepackage{graphicx}
\usepackage[labelfont=bf]{subcaption}

\usepackage{tcolorbox}% http://ctan.org/pkg/tcolorbox
\definecolor{mycolor}{rgb}{0,0,0}% Rule colour
\makeatletter
\newcommand{\mybox}[1]{%
	\setbox0=\hbox{#1}%
	\setlength{\@tempdima}{\dimexpr\wd0+13pt}%
	\begin{tcolorbox}[colframe=mycolor,boxrule=0.5pt,arc=4pt,
		left=6pt,right=6pt,top=6pt,bottom=6pt,boxsep=0pt,width=\@tempdima]
		#1
	\end{tcolorbox}
}
\makeatother
\usepackage[resetlabels,labeled]{multibib}
\newcites{A}{Appendix}
\usepackage[num]{abntex2cite}
\citebrackets[]
%----------------------------------------------------------------------------------------
%	COLUMNS
%----------------------------------------------------------------------------------------
\setlength{\columnsep}{0.55cm} % Distance between the two columns of text
\setlength{\fboxrule}{0.75pt} % Width of the border around the abstract
%----------------------------------------------------------------------------------------
%	COLORS
%----------------------------------------------------------------------------------------
\definecolor{color1}{RGB}{0,0,90}
\definecolor{color2}{RGB}{0,20,20}
%----------------------------------------------------------------------------------------
%	HYPERLINKS
%----------------------------------------------------------------------------------------
\usepackage{hyperref} % Required for hyperlinks
\hypersetup{hidelinks,colorlinks,breaklinks=true,urlcolor=color2,citecolor=color1,linkcolor=color1,bookmarksopen=false,pdftitle={Title},pdfauthor={Author}}
%----------------------------------------------------------------------------------------
%	ARTICLE INFORMATION
%----------------------------------------------------------------------------------------
\JournalInfo{\color{gray}\normalsize\sffamily\bfseries Revista de Informática Teórica e Aplicada - RITA - ISSN 2175-2745\\ Vol.~XX, Num.~XX~(2018)~11-XX} % Journal information
\Archive{\mybox{RESEARCH ARTICLE}} % Type of Article: Researh, Review, Tutorial, Study Case

\PaperTitle{Symmetric Peer-to-Peer Applications}
\ShortTitle{Symmetric Peer-to-Peer Applications}

\Titulo{Aplicações Peer-to-Peer Simétricas}

\PaperVol{XX} % Volume 
\PaperNum{X} % Number 
\PaperAno{YYYY} % Year of Publication

\Authors{Francisco Sant'Anna\textsuperscript{1}*}
\affiliation{\textsuperscript{1}\textit{Department of Computer Science, Rio de Janeiro State University (UERJ), Brazil}}
\affiliation{*\textbf{Corresponding author}: francisco@ime.uerj.br}

\Keywords{collaboration --- determinism --- peer-to-peer --- time machine }

\newcommand{\keywordname}{Keywords} % Defines the keywords heading name

\PalavrasChave{colaboração --- determinismo --- peer-to-peer --- máquina do tempo} 
\Doi{http://dx.doi.org/10.22456/2175-2745.XXXX} % DOI
\DateR{dd/mm/yyyy} % Received
\DateA{dd/mm/yyyy} % Accepted

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\Abstract{
In real-time networked applications, such as shared documents, users can
interact remotely and yet share the exact same experience as if they were in a
single machine.
%
In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications} in which decentralized instances can trigger asynchronous events
and yet conform to identical behavior.
%
Peers are allowed to join and leave the network at any time.
Application developers must adhere to a restricted API, which is deterministic,
stateless, and only supports pre-allocated memory.
%
The middleware is responsible for synchronizing the events in a global shared
timeline across the network.
Also, based on memory snapshots and deterministic simulation, a ``time machine''
can rollback conflicting peers to resynchronize their state.
%
We show that the middleware can handle applications with over 20 peers in a
heterogeneous topology under high churn.
For instance, in a simulation scenario of 25 events per minute with a network
latency of 50ms, the peers need to roll back and resynchronize only 3\% of
the time.
}

\Resumo{
Em aplicações de tempo real em rede, tais como documentos compartilhados, os
usuários podem interagir remotamente e ainda obterem a mesma experiência, como
se estivessem em uma única máquina.
%
Neste trabalho, nós propomos um middleware para \emph{aplicações peer-to-peer
simétricas} nas quais instâncias descentralizadas podem disparar eventos
assíncronos e ainda obterem comportamento idêntico.
%
Os pares podem entrar e sair da rede a qualquer momento.
Os desenvoledores de aplicações devem aderir a uma API restritiva, que deve ser
determinística, sem estados internos, e também com alocação prévia de memória.
%
O middleware é responsável por sincronizar os eventos em uma linha de tempo
global e compartilhada em toda a rede.
Além disso, com base em cópias da memória e simulação determinística, uma
``máquina do tempo'' pode retroceder pares em conflito para ressincronizar o
seu estado.
%
Nós mostramos que o middleware suporta aplicações com mais de 20 pares em uma
topologia heterogênea e sob alta taxa de desconexões.
Como exemplo, em uma simulação com 25 eventos por segundo em uma rede com
latência de 50ms, os pares precisam retroceder e ressincronizar somente 3\% do
tempo.
}

%----------------------------------------------------------------------------------------

@@

%\setcounter{page}{11}

\flushbottom % Makes all text pages the same height
\maketitle % Print the title and abstract box
\thispagestyle{empty} % Removes page numbering from the first page
%\tableofcontents % Print the contents section

\section{Introduction} % The \section*{} command stops section numbering
\label{sec.introduction}

@@

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Peers in the application form a dynamic network graph and communicate only
with direct neighbours, as in typical unstructured peer-to-peer
networks~\cite{p2p.survey}.
Events are flooded in the graph and are triggered locally with a small delay
to compensate the network latency.
To deal with events received out of order or too late, a time machine can
rollback peers locally to a previous state and reapply events in order and
in time.
Our main contribution is the design of a middleware to ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.
We target small peer-to-peer networks, in which nodes are only a few hops away
from each other.
This ensures that events can span the whole network in a reasonable time to
preserve the real-time behavior of applications.
We also assume that peers are non-malicious in the sense that they do not
generate erroneous events.

@@

In Section~\ref{sec.related}, we revisit existing solutions for symmetric
distributed applications and software time machines.
In Section~\ref{sec.tml}, we detail the architecture of our middleware, its
programming API, and how it orchestrates peers.
%In Section~\ref{sec.app}, ...
In Section~\ref{sec.eval}, we evaluate our design under a number of scenarios.
In Section~\ref{sec.conclusion}, we conclude this work.

@@

This section first revisits existing solutions for symmetric distributed
applications, namely Croquet~\cite{croquet,croquet.new}, GALS~\cite{gals}, and
CRDTs~\cite{crdts}.
Then, we discuss software time machines that are built for either local or
networked applications.

\subsection{Symmetric Distributed Applications}
\label{sec.related.sym}

@@

Even though existing solutions aim to provide symmetric behavior for
distributed applications, they diverge in three key aspects as follows:

\textbf{a.} How the network is organized.

\textbf{b.} How global time is determined.

\textbf{c.} How events are propagated and applied.

Figure~\ref{fig.table} compares three selected works regarding these aspects,
which we discuss and contrast with our approach next.
At the end of this section, we also provide initial details on how our
middleware works.

Croquet%
\footnote{Croquet.io: https://croquet.io/}~\cite{croquet} guarantees
bit-identical real-time behavior for users in collaborative distributed
environments.
%
As detailed in Figure~\ref{fig.algos}, a centralized \emph{reflector} issues
periodic ticks, such that all nodes in the network advance together according
to a synchronized global shared clock.
If a user generates an event, it is sent back to the reflector, which
broadcasts the event in the next tick, which all nodes apply in sync.
%
Croquet takes periodic snapshots of the whole application state in order to
support late joins to a running session.
The new node just needs to restore the latest snapshot and simulate the
remaining events to reach the current running state.
%
As indicated in Figure~\ref{fig.table}, Croquet relies on a centralized
network (a), in which nodes advance in sync (b), and in which event outcomes
depend on the central server to be online (c).
%
Note how the central server is the only participant with knowledge about
clients, which never communicate directly among themselves.
This implies that the server must never fail, and also that periodic full
broadcasts are the only possibility of spanning the whole network.
In this work, we argue that in a peer-to-peer alternative any node can fail,
and that broadcasts can be replaced by optimistic flooding.

GALS~\cite{gals} shares the same goals with Croquet, but with some tradeoffs,
mostly favoring clients with slow connections.
As detailed in Figure~\ref{fig.algos}, instead of advancing ticks in sync with
the server, clients have their own independent local clocks.
Event generation requires two round trips (\emph{when → ask → ack → tick}) and is
delayed dynamically according to the slowest client.
%
On the one hand, ticks do not generate any traffic and clients have smooth
frame transitions, even those with poor connections.
On the other hand, events take longer to apply and clients may experience
occasional freezes.
The syncing protocol also requires extra bookkeeping to deal with clock drifts
and client disconnections~\cite{gals}.
%
As indicated in Figure~\ref{fig.table}, GALS also relies on a centralized
network (a), but in which nodes advance time independently (b), and in which
event outcomes still depend on the central server to be online (c).

@@

In this work, as indicated in Figure~\ref{fig.table}, our goal is to support
peer-to-peer real-time symmetric execution (a), while still tolerating
out-of-order (b) and offline (c) event generation.
%
As detailed in Figure~\ref{fig.algos}, peers have independent timelines and
can communicate events directly to each other.
Note, that it is not necessary that all peers communicate directly with all
other peers, but only indirectly via unstructured flooding.
Events are delayed to be able to reach other peers in time.
In the case a deadline is missed, the peer rolls back in time and then fast
forwards execution until it catches real-time behavior again.
%
Like Croquet and GALS, peers compute events in sequence using deterministic and
stateless calls only.

Finally, note that classical distributed consensus mechanisms focus on fault
tolerance~\cite{lamport}, but not on real-time symmetric behavior.
These algorithms typically rely on leaders~\cite{consensus.paxos.raft}, which
require an election period that does not cope with our symmetric real-time
requirements.
In this sense, these algorithms could be considered to decentralize servers in
Croquet or GALS, but not to decentralize symmetric peers with high churn rates,
as we propose in this work.
In contrast, our proposal requires a time machine to reapply reordered
operations, since we cannot guarantee strong consistency~\cite{p2p.sec}.

\subsection{Software Time Machines}
\label{sec.related.time}

%We now discuss related work on decentralized and centralized time machines.

A software time machine is a mechanism that allows to regress or advance the
execution of an application.
%
Time travelling is a established research topic for debugging
tools~\cite{tml.review}.
In this context, it is possible to stop a program after a failure and revert
its execution in order to understand the bug causes.
%
In our context of real-time symmetric applications, time travelling is
essential to re-synchronize the execution of mirrored instances that diverge
due to asynchronous events.

Fundamentally, it is not possible to reverse the execution of most programs.
The reason is because even trivial operations, such as `printf` and `abs` are
not invertible.
Therefore, time machines rely on techniques that either \emph{record} the
entire application states in a log, or \emph{reconstruct} them from
deterministic inputs~\cite{tml.review}.
%
In this work, we mix both techniques by taking periodic memory snapshots and
simulating deterministic inputs in between them.

Regarding decentralized machines, Fusion%
\footnote{Fusion: \url{https://doc.photonengine.com/en-us/fusion}}
is a state synchronization networking library for the Unity game engine.
It provides networked tick-based simulation in two modes: client-server
(\emph{hosted mode}) and peer-to-peer (\emph{shared mode}).
The hosted mode is based on continuous memory snapshots, which allows for
full rollbacks when clients diverge.
As discussed in Section~\ref{sec.related.sym}, this approach is similar to how
Croquet handles late joins to a running session.
The shared mode is less powerful and only supports eventual consistency among
clients.
We presume that the reason is because continuous memory snapshots would be too
costly without a central server.

Regarding single-user local machines, the video game Braid%
\footnote{Braid: \url{http://braid-game.com/}}
is designed on the assumption that players have unrestricted rollbacks as part
of the game mechanics.
Like Fusion, the implementation is based on continuous memory snapshots,
instead of deterministic simulation, as we propose in this work.
%
Schuster~\cite{tml.js} proposes a simple API for time travelling in
event-driven JavaScript:
    \code{init()} to return an initial state;
    \code{handle(evt,mem)} to process events and modify state; and
    \code{render(mem)} to output the current state.
As we describe next, we use a similar approach by separating state modification
from rendering~\cite{tml.alive}.

@@

We guide our discussion through the didactic example of Figure~\ref{fig.move},
in which multiple users use the arrow keys to control the direction of a
moving rectangle.
%
The application is symmetric in the sense that if any user presses a key, a
corresponding event propagates in the network, and all peers apply it
simultaneously, as if they were a single mirrored machine.
At any moment, if a user pauses the application, all peers are guaranteed to
pause exactly at the same frame with the rectangle being ``bit-identically'' at
the same position.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{middleware}
  \caption{
    Applications generate asynchronous events and communicate through an API
    with the middleware.
    The middleware orchestrates the peers, synchronizing their ticks locally
    and disseminating events with a $\Delta$ delay.
    \label{fig.middleware}
  }
\end{figure}

In Figure~\ref{fig.middleware}, a peer is depicted as an application that uses
an API to communicate with the middleware.
The application itself is not aware of the network and has no direct access to
other peers, which communicate transparently through the middleware.
The application generates asynchronous events that go through the middleware,
which synchronizes them with a timestamp with a $\Delta$ delay scheduled to the
future, such that all peers can (optimistically) satisfy.
The middleware controls the execution of the application by issuing ticks and
synchronized events.

The peers form a dynamic unstructured peer-to-peer network~\cite{p2p.survey},
and communicate mostly indirectly to each other.
Events are flooded in the network graph via gossiping, i.e.: when a peer
generates an event, it communicates to its neighbours, which communicate to
their neighbours, and so on.
Note that all peers execute the exactly same application and middleware, with
no differences with respect to their roles and physical resources.
Therefore, we can describe our peer-to-peer network as follows:
    (a) all peers have the same role and run the same software;
    (b) any peer can join or leave at any time;
    (c) events are only communicated with direct neighbours.

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
As described in the Introduction, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate delta $\Delta$ so that they can reach other peers in time to be
applied in sync.
%We first assume a predetermined delay for the sake of simplicity.
Nevertheless, a peer ahead of time may receive an event that should have been
applied in the past, even considering the delay.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
As we detail further, rolling back a peer requires to simulate the whole
application execution from the beginning.
As an optimization, the middleware takes periodic snapshots locally at all
peers to avoid full simulation.
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because applications are launched locally at different times, and
because internal clocks may diverge over time (e.g., from rollbacks or timer
polling inaccuracies).
The middleware assumes the maximum time among all peers to be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers disseminate \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
\item \textbf{Peer Churn:}
Considering that we target peer-to-peer networks, it is important that the
middleware can handle arrival and departure of peers seamlessly.
Note that high churn may lead to intermittent network partitions.
Given our unstructured approach, nothing needs to be done when a peer leaves
the network.
However, when a peer joins the network, the middleware needs to ensure that it
receives from other peer all events that ever happened, in order.
This is trivial since all peers already need to hold the full event history (as
described in item (a)), including their absolute timestamps in the shared
timeline.
In addition, a peer that rejoins the network recovers from a past snapshot it
already holds, thus avoiding full simulation.
\end{enumerate}

@@

When an application generates events, they need to propagate in the network
so that all peers behave the same.
Note that application events may (or may not) differ from low-level local
system events.
%which may or may not be mapped 1-to-1 in the callback \code{cb\_evt}.
%
The middleware predefines two events for all applications:
    \code{P2P\_EVT\_START} represents the beginning of the simulation, and
    \code{P2P\_EVT\_TICK} is generated every frame.
The other application-specific events must be declared by the programmer in an
enumeration.
In our example, we define a new event \code{P2P\_EVT\_KEY} to represent key
presses:

@@

The callback \code{cb\_evt} is called by the middleware in real time, whenever
a local SDL event occurs.
This callback has two goals:
    (a) map and propagate application events, and
    (b) provide instant feedback to the user.
%As mentioned previously,
Not all local low-level events need to correspond to application events in the
network.
The callback is allowed to perform side effects, such as modifying the network
topology, or exhibiting instant feedback on the screen.
In our example, we only generate application events for the 5 key events of
interest, but we also exhibit the pressed key in real time as a visual
feedback:

@@

\subsubsection{Event Dissemination}
\label{sec.tml.middleware.events}
